---
title: "Donnermeyer_Homework_2"
author: "Sam Donnermeyer"
date: "9/11/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load Packages
library(tidyverse)
library(ggplot2)
library(ggeasy)
library(scales)
library(grid)
library(gridExtra)
library(patchwork)
library(utils)

# Formating numbers made through in-line code
inline_hook <- function(x) {
  if (is.numeric(x)) {
    format(x, digits = 2)
  } else x
}
knitr::knit_hooks$set(inline = inline_hook)
```

## Hand-In Homework #2

```{r data, include=FALSE}
# Read in data 
dalmatian <- read_csv("dalmatian.csv")
epa_tccb <- read_csv("EPAtccb.csv")
```

```{r tire_variables, include=FALSE}
# Set Constants for the Questions
mean_tire = 32000
std_tire = 2500
qnorm(.04, mean_tire, std_tire)
```

### 1 - Problem #54, pg. 156

A tire manufacturer believes that the tread life of its snow tires can be described by a Normal model with a mean of 32,000 miles and a standard deviation of 2,500 miles. 

  a) If you buy one of these tires, would it be reasonable for you to hope it will last 40,000 miles? Explain.
  
  No, it would be totally unreasonable for you to hope that the tires will last 40,000 miles because only about `r percent(1- pnorm(40000, mean_tire, std_tire), accuracy = 0.1)` tires will last until 40,000 miles of use. 
    
  b) Approximately what fraction of these tires can be expected to last less than 30,000 miles? 
  
  According to the Normal model, approximately `r percent(pnorm(30000, mean_tire, std_tire), accuracy = 0.1)` of tires will last less than 30,000 miles. 
    
  c) Approximately what fraction of these tires can be expected to last between 30,000 and 35,000 miles? 
  
  According to the Normal model, approximately `r percent((pnorm(35000, mean_tire, std_tire) - pnorm(30000, mean_tire, std_tire)), accuracy = 0.1)` of tires would last between 30,000 and 35,000 miles. 
    
  d) Estimate the IQR of the tread lives. 
  
  The IQR of the tire's tread life is approximately `r round(qnorm(.75, mean_tire, std_tire) - qnorm(.25, mean_tire, std_tire), 1)` miles which is from `r qnorm(.75, mean_tire, std_tire)` - `r qnorm(.25, mean_tire, std_tire)`. 
    
  e) In planning a marketing strategy, a local tire dealer wants to offer a refund to any customer whose tires fail to last a certain number of miles. However, the dealer does not want to take too big a risk. If the dealer is willing to give refunds to no more than 1 of every 25 customers, for what mileage can he guarantee these tires to last? 

  If the local tire dealer wants to refund no more than 1 of every 25 customers, the tire mileage must be guaranteed to at a minimum `r qnorm(.04, mean_tire, std_tire)` miles. 


### 2 - Hearing Impairment in Dalmatian Dogs
The relationship between eye color and degree of impairment: deaf, unilateral hearing (one ear), bilateral hearing (both ears). 

#### Two-Way Table of the counts with marginal totals
``` {r dalmatian_TwoWayTable}
# Create two-way table of dalmatian tibble 
dal_table = addmargins(xtabs(Count~Hearing+EyeColor,data=dalmatian))
dal_table
```

#### A two-way table showing the conditional distribution of the hearing variable given each eye color.
``` {r dalmatian_TwoWayConditional}
# Conditional distribution of hearing variable given each eye color 
dal_condit = round(prop.table(xtabs(Count~Hearing+EyeColor,data=dalmatian),2),3)
dal_condit
```

#### Segmented bar chart comparing the conditional distributions in the previous table.
``` {r dalmatian_SegmentedBarChart}
# Segmented bar chart
ggplot(dalmatian,aes(x=EyeColor,y=Count,fill=Hearing)) + 
  geom_col(position="fill") +
  # Change the Y-axis to percentage
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  ylab("Proportion") + 
  # Change Fill colors 
  scale_fill_manual(values = c("darkslategray3", "cadetblue", "darkslategrey")) +
  # Theme
  theme(
    # hide panel borders and remove grid lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_blank(), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    # change axis titles 
    axis.line = element_line(color = "black"),
    axis.title = element_text(size=9),
    # Change the title orientation 
    plot.title = element_text(hjust = 0.5)
  )
``` 

#### Association between Hearing and Eye Color within Dalmatians

Generally, Dalmatians with blue eyes, the greater the incidence of unilateral hearing and deafness. In the Dalmatians with blue eyes, 32.3% have unilateral hearing and 18.4% are deaf compared to 20.7% and 6.8% respectively for those with blue eyes. 
