---
title: "Donnermeyer_Homework_1"
author: "Sam Donnermeyer"
date: "8/27/2020"
output: html_document
---

```{r setup, include=FALSE}
# Load Packages
library(tidyverse)
library(ggplot2)
library(ggeasy)
library(scales)
library(grid)
library(gridExtra)
library(patchwork)
```


```{r,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hand-In Homework #1

```{r, include=FALSE}
growth_rates <- read_csv("GrowthRates.csv")
head(growth_rates, 3)
```

### Create Histogram and Density Plot

```{r, echo = TRUE}
## 2a ## 
# Create Histogram of Rate2015to20
histogram_plot = ggplot(growth_rates, aes(Rate2015to20)) +
  geom_histogram(binwidth = 3) + 
  # Add x label
  xlab("Estimated Population Growth Rate (%) 2015-20") +
  # add y label
  ylab("Count of Nations") +
  # Add title to the chart
  ggtitle("Growth Rates During 2015 - 2020") +
  # Theme
  scale_colour_viridis_d() +
  theme(
    # hide panel borders and remove grid lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_blank(), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    # change axis titles 
    axis.line = element_line(color = "black"), 
    axis.title = element_text(size=9),
    # Change the title orientation 
    plot.title = element_text(hjust = 0.5)
  )

# Create Density plot of Rate2015to20
density_plot = ggplot(growth_rates, aes(Rate2015to20)) + 
  # Density plot 
  geom_density(color = "black", fill = "light grey", alpha = .4) +
  # Add x label 
  xlab("Estimated Population Growth Rate (%) 2015-20") +
  # Add y label
  ylab("Proportion of Nations") +
  # Add title to the chart
  ggtitle("Growth Rates During 2015 - 2020") +
  # Themes
  scale_colour_viridis_d() + 
  #theme_minimal() + 
  # Change the Y-axis to percentage
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  # Theme
  theme(
    # hide panel borders and remove grid lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_blank(), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    # change axis titles 
    axis.line = element_line(color = "black"),
    axis.title = element_text(size=9),
    # Change the title orientation 
    plot.title = element_text(hjust = 0.5)
  )

histogram_plot + density_plot 
```

The distribution of estimated population growth rate for 235 given countries and territories is mostly unimodal with a large peak around five percent. The growth rates are generally evenly distributed, except for a small increase in nations with about a thirteen percent growth rate increase. There is a single country that resides outside of the distribution at negative fifteen percent. 

The mean of the growth rate is `r mean(growth_rates$Rate2015to20)`, which is the best measure of center in this generally unimodal distribution. The median of the growth rate is `r median(growth_rates$Rate2015to20)`. The IQR of the growth rate distribution is (`r quantile(growth_rates$Rate2015to20, 0.25)`, `r quantile(growth_rates$Rate2015to20, 0.75)`) which reveals the spread of the distribution accurately.

### Create Boxplots by Continent

```{r, echo = TRUE}
## 2d ##
# Create side-by-side boxplots of growth rates for the 6 continents 
boxplot = ggplot(growth_rates, aes(Continent, Rate2015to20)) +
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE) +
  # Add x label 
  xlab("Continent") +
  # Add y label
  ylab("Estimated Population Growth Rate (%) 2015-20") +
  # Add title to the chart
  ggtitle("Estimated Growth Rates During 2015 - 2020") +
  # Themes
  scale_colour_viridis_d() + 
  # Theme
  theme(
    # hide panel borders and remove grid lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_blank(), 
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    # change axis titles 
    axis.line = element_line(color = "black"),
    axis.title = element_text(size=9),
    # Change the title orientation 
    plot.title = element_text(hjust = 0.5)
  )

boxplot
```

The continent that tended to have the largest growth rate is Africa and Europe had tended to have the lowest. Africa's estimated growth rates contained a lot of variation and South America had the least variation in the growth rate. 
